Neutral models of quantitative trait evolution are important for establishing a
baseline against which to test for selection. \citet{Schraiber2015} recently
analyzed a neutral model of trait evolution that made few assumptions about the
number of loci potentially affecting the trait and the distribution of
mutational effects at these loci. However, they only derived results for
constant-size, panmictic populations. We extend their results to populations
with arbitrary distribution of coalescent times and therefore varying
demographies and population structures. As their key result,
\citet{Schraiber2015} derived the characteristic function of the distribution of
trait values in a sample. In this paper, we instead work with the moment
generating function, but the two approaches are interchangeable as long as the
mgf for the distribution of mutational effects exists (which it will if the
moments of the mutational distribution are all finite). Our main result, given
in equation \eqref{eq:sub}, shows that the characteristic function
from \citet{Schraiber2015} is a special outcome of a general procedure whereby
the moment generation function for a trait distribution can be obtained by
making a simple substitution into the moment generating function for a
distribution over genealogies. The moment generating functions for many
demographies of interest and for sample sizes above three are sufficiently
complex that solving for them is impractical \citep{Lohse2011}. However,
progress can still be made by using Taylor expansions to write moments of the
trait distribution in terms of moments of the genealogical and mutational
distributions.

This result extends previous work using coalescent theory to investigate neutral
models of quantitative traits \citep{Whitlock1999,Schraiber2015}. Ours is the
most general model yet analyzed. As a natural first step, we show that the
infinitesimal limit suggested by \citet{Fisher1918} leads to a model where
phenotypes are normally distributed as the number of loci becomes large and the
variance of effect sizes becomes small. In the limiting distribution, the
variance of the difference in trait values between two individual is
proportional to the expected pairwise coalescent time between them, and the
covariance between a pair of differences more generally is $\Cov[Y_a - Y_b,Y_c -
Y_d] \propto \E[\mathcal{T}_{a,d}] + \E[\mathcal{T}_{b,c}] -
\E[\mathcal{T}_{a,c}] - \E[\mathcal{T}_{b,d}]$. The resulting covariance matrix
completely specifies the neutral distribution under the infinitesimal model.
This is similar to classic models in evolutionary quantitative genetics
considering the neutral divergence of trait values after population splits
\citep{Lande1976,Lynch1989} but holds regardless of the precise details of
population structure and history. \citet{Schraiber2015} derive essentially the
same distribution using a central limit theorem argument. 

\citet{Mendes2018} point out several problems caused by incomplete
lineage sorting when species split times are used to generate the covariance
matrix in phylogenetic comparative methods. Since the normal model arising in
the infinitesimal limit is explicitly based on the coalescent, it is not subject
to these problems. A matrix based on pairwise coalescence times rather than a
species tree based on population split times takes into account the effects of
all lineages at causal loci, even those that do not follow the species topology.
The covariances specified by equation \eqref{eq:normcov} could also be used to
generate within-species contrasts in a similar manner to the method suggested
by \citet{Felsenstein2002}.

We compared sparse traits to the normal model by calculating how the first four
expected central moments different from those expected under normality. This
showed how demography and genetics separately influence the expected deviation
from normality. For a fixed expected trait sparsity, population growth produces
greater deviations in the fourth central moment while population bottlenecks
produce lower deviations (Figures
\ref{fig:Qexp}). However, for realistic demographic scenarios,
we find that the effects attributable to demography are small (Figure
\ref{fig:afeucomp}). We only analyzed cases where the individuals in the
population were exchangeable, but adding population structure would increase
deviations from normality as drifting trait means between subpopulations will
yield a multimodal distribution.

We next apply the above theory to two simple problems where a coalescent
perspective on the neutral distribution of a quantitative trait provides useful
intuition. The first of these is the question of the appropriate null
distribution for $Q_{ST}$ at the population level. We show how the null
distribution under the normal model can be easily simulated from by taking
advantage of the theory presented here, providing a much better approximation
when populations are correlated than previous approaches \citep{Whitlock2009}
(Figure \ref{fig:qst_deme}). In the second we show that the shape of the
mutational distribution is largely confounded with the number of loci affecting
the trait, with only one compound parameter identifiable. This makes it unlikely
that mutational parameters could be inferred from trait values sampled from a
population.

Even though we have broadened the model space for neutral traits, many features
of real populations have not yet been incorporated. Linkage between loci is a
particular concern as there is substantial linkage disequilibrium between causal
loci \citep{Bulik-Sullivan2015}. \citet{Lohse2011} derived the form of the
moment generating function for linked loci and future work will attempt to
incorporate this using equation \eqref{eq:sub}. In particular, it will be
important for future work to show how this affects the distribution in the
infinitesimal limit. Diploidy, dominance, and epistasis have also been ignored
thus far. The qualitative effects described here should hold under diploidy, but
having trait values within individuals summed over loci from two copies of the
genome will decrease deviations from normality. Dominance will also tend to
produce a normal distribution as the effects are independent between loci, but
future work is needed to examine how this interacts with the distribution of
genealogies to affect the trait distribution.

\citet{Barton2017} recently performed a deep mathematical investigation of a
more formal ``infinitesimal model''. They proved conditions under which the
trait values of offspring within a family are normally distributed with
covariance matrix independent of the parental trait values conditional on the
pedigree. Interestingly, they found normality still holds under pairwise
epistasis if it is not too extreme. It may therefore be possible to include
epistasis in the infinitesimal limit considered here and thus study how
epistasis affects neutral trait divergence. In a commentary on the
\citet{Barton2017} paper, \citet{Turelli2017} noted that at least three
different sorts of infinitesimal models have been used. The infinitesimal limit
considered here fits all three: (1) the limit corresponds to Fisher's notion of
a large number of Mendelian factors, each making a small contribution
\citep{Fisher1918}; (2) the descendants within a family will have a Gaussian
distribution of trait values, as shown rigorously by \citet{Barton2017}; and
(3) the distribution of trait values in a panmictic population will be Gaussian.

Although GWAS of many traits have shown them to be controlled by large numbers
of loci \citep{Boyle2017}, this will not necessarily be the case for every trait
of interest to biologists. It has been suggested, for instance, that gene
expression levels have a sparse genetic architecture \citep{Wheeler2016}. Since
there is much interest in testing whether natural selection has acted on gene
expression levels \citep{Whitehead2006,Gilad2006,Yang2017}, well-calibrated
goodness-of-fit tests will need to take into account the complications that
arise when trait distributions deviate from normality \citep{Khaitovich2005}.
Direct measurements of mutational distributions \citep{Gruber2012,Metzger2016}
could aid in such a calibration. Finally, equation \eqref{eq:emoms} suggests a
means to determine whether sparsity impacts trait distributions by examining
populations with different demographic histories. Populations with a greater
$\mathbbm{T}_{3,3}$ to $\mathbbm{T}_{2,2}$ ratio are also expected to show a
greater $M_3$ to $M_2$ ratio and an equivalent relationship would be expected
for the fourth moment. Because gene expression studies generally measure a large
number of traits, observing such a trend on average could be a sign of
deviations from normality. Generally, more work is needed to develop neutrality
tests that are robust to the details of genetics and population history, and to
investigate whether any information about mutation can be uncovered by
statistical models that share information across multiple measured traits. 

%%% Local Variables:
%%% TeX-master: "quant_gen_manu.tex"
%%% End:
